\textsc{Ferret Miner}: A Process Mining Case Study${}^\dagger$\thanks{${}^\dagger$~Github: \texttt{https://github.com/Martin-Chivo/CMPE255-Team8}; Google Drive: \texttt{shorturl.at/cgFIP}}

### 2022-12-14 paper/paper.tex
Minor edits.

\begin{abstract}
This project is a process mining case study designed to
identify patterns and
delays that obstruct the efficient flow of process
activities. Specifically, our
objective is to analyze event data logs from
the travel reimbursement process at Eindhoven University of Technology
(ETU/c), identify bottlenecks in the process, and
seek improvement to its efficiency.
We prepare and filter the event data, and then generate simplified and
representative process models to locate
the most time-consuming and impactful critical tasks.
Our analysis suggests that
improvement in the supervisor approval and payment handling activities 
could significantly
reduce the time processing the travel reimbursement.
\end{abstract}

@@

Nowadays, most organizations use information systems to support
the execution of their business processes, generating significant
amounts of data.
Mainstream information systems, such as Enterprise Resource
Planning (\textsc{erp}), Work on Management Systems (\texttt{wms}), and Customer
Relationship Management (\textsc{crm}), have become vital tools behind the
operation of almost 
any company process\cite{Tuto2022}. All of these platforms
offer rich event logging capabilities, in which
an event log is
basically a record of an activity performed with a timestamp embedded
\cite{Proc2022}.

There are plenty of oppportunities, through
proper supervision of business behavior, to identify and adjust
tasks that are generating problems for the business processes
\cite{LeAl1994}.
One important technique towards such goal is process mining, a
specialized data analysis technique that reveals core business behavior
indicated by activity logs.
Process mining would start with these event logs, extracting important
information including the chronology of the activities
 and their execution time registered in each log. Representative models
would be built, or discovered automatically,
and then be used to detect errors and deficiencies in the process,
and to provide insight and suggestions for process enhancement.
Process mining is most closely related to data mining 
amongst business intelligence disciplines. However,
data mining concentrates on finding relationships among data,
whereas process mining also looks at data from a systematic
perspective.

@@

In our process mining case study, we investigated
a dataset with
real-life event log data from the travel reimbursement process at
Eindhoven University of Technology (ETU/e). Through analyzing the event
logs and mining the business process underlying the events, we seek to
identify and suggest remedy to address deficiencies in the reimbursement
process.
The event logs provide a detailed record of the travel reimbursement process,
including important attributes such as identification, resource,
timestamp, permits, payment and budgetting information, and event labels.
Students or staff would submit their reimbursement application,
provide suppport documentation (if traveling internationally), wait for
reimbursement decision and subsequent payment, and possibly deal with
rejection and resubmission. All of these information is captured in the
event log.
Collectively, these information provided a substantial amount of detail
into the process. 
The data was obtained from 
the annual \textsc{bpi} Challenge as part of
the 2020 International Conference on Process Mining \cite{BPI2020}.

Our approach to process mining is based on the work by
Wil van der Aalst in his seminal paper titled simply
``Process Mining'' \cite{van2012},
where process mining guidelines are detailed.
A process mining implementation generally
covers three main issues: (1) discovery, (2) conformance,
and (3) enhancement.
Our project focuses on discovery since it is the area
that requires more effort and it is the most deterministic for the
success of the analysis. We will also take a cursory
look at conformance as both an aide to discovery and future work.

@@

Publications on practical application of process mining are scarce,
and our research paper aims to add support to this section. Our goal is
to demonstrate the practicality of process mining analysis
in practice, in application to the real-life travel reimbursement problem
as detailed earlier.

Furthermore, we are very interested in the evaluation and application
of a relatively new process mining technique based on
\textsc{pm4py}, a Python process mining library
developed by Berti et al.~\cite{BeZe2019}.
According to the authors,
\textsc{pm4py} is a modern, open-source, and emerging platform developed
specifically for the data science community. Prior to the turn of the
millenium,
there were hardly any tooling for process mining, but
since then a slew of software tools, both open-source
(\textsc{prom} and \textsc{apromore}) and commercial
(\textsc{disco}, \textsc{celonis}, and \textsc{processgold}),
have been developed. Yet, many were proprietary, hard to extend,
or hidden behind graphical user interfaces.
The mindset of process mining
tooling contrasted greatly with the modern philosophy of data science R\&D,
which put significant emphasis on openness and extensibility.
\textsc{pm4py} was developed to bridge the gap
between process mining and modern data science
capabilities including
\textsc{pandas}, \textsc{numpy}, \textsc{scikit-learn}.
We whole-heartedly
agree with the assessment of the authors, and we look further at opportunities
that the \textsc{pm4py} approach brought forward, including
the adoption of modern and rapidly evolving machine learning
functionalities to process mining.

@@

The document is structured as follows: Sections \ref{section-intro} and
\ref{section-survey}
offer an introduction to process mining and a literature survey.
Sections \ref{section-technical}, \ref{section-methodology} and
\ref{section-models} describe the technical details of our methodology, as
well as the results of our investigation. We will end the paper with
future work in Section \ref{section-future} and conclusion in
\ref{section-conclusion}.

@@

\begin{figure}[htbp]
\centerline{\includegraphics[width=0.35\textwidth]{images/image2a.png}}
\caption{General outline of process mining approach.}
\label{fig-outline}
\end{figure}

@@

Many distinct data analysis approaches have been developed to discover
processes from event logs. However, \textsc{petri net}s\footnote{and the
equivalent BPMN models and process trees} have become a
widely adopted technique over other methods \cite{Cook1998} due to their
intuitiveness, mathematical soundness, and ease of analysis.
Another approach utilizes Unified
Modeling Language (\textsc{uml}) diagrams \cite{Sait2019}.
In 2004, van der Wilst et al.~described the 
a break-through algorithm in process mining
called \textsc{alpha miner}\cite{van2004} which
is capable of depicting event data logs into
work-flow diagrams, which can be translated to \textsc{petri net}s.
The success of \textsc{alpha miner} has inspired many follow-up publications,
including a popular model
called \textsc{flexible heuristic miner} \cite{Weij2006}.
One main characteristic of this
algorithm is its high level of configurability and flexibility.
In order to deal with the
massive number of events that actual information systems generate,
Burattin et al.~\cite{Bura2014} developed another extension to
\textsc{heuristic miner} based on
lossy counting, as well as a sliding window that effectively analyzes
streams of real datasets.

@@

Most process mining publications are dedicated mainly to the enhancement
of novel methodologies and algorithms with significant emphasis on the
control-flow discovery area \cite{van2004}. However, the public services area
has been fruitful ground
for process mining deployment research. For instance, a
genetic miner approach was tested by de Madeiros et al.~\cite{deMe2007}
in Dutch municipalities. Additionally, van del Aalst et al.~\cite{RoDJ2009}
applied an analysis of organizations in another municipality process.
Rozinat et al.~\cite{RoMa2009} evaluated process models in two different case
studies on the public sector.
Nonetheless, all these research
applications were focused on validation and assessment,
rather than to improve the process.

@@

We shall now briefly describe our approach to finding
where the process bottlenecks are. We would start with a thorough analysis and
cleaning of the event logs, and then
build models that best capture the behavior of the
process, making the proper tradeoff
between robustness (does the model capture the intent of the process),
and coverage (what fraction of the use cases are captured).
 Then, we will use the
chosen models to answer the questions we have in mind.

We shall note that process mining is a relatively new techniques, and
as such we spent a disproportional amount of time ramping up the project.
While we fully believe in the full potential of \textsc{pm4py} due to
its focus on algorithmic discovery and availability
on Python, it is
essentially still an academic software in its nascent stage, and 
hence suffers from instability. There is still plenty of room for
improvement, such as a lack of control in visualization and
presentation, scanty documentation, and signs of immaturity in API design.

@@

After a thorough review of what our tasks entail, we decided to focus
exclusively on the domestic and international travel event logs in order
to identify the process bottlenecks.  Both event logs
were provided to us in \textsc{xes} format,
an \textsc{xml}-based standard for event logs \cite{XES2021}, which can
be easily parsed using \textsc{pm4py}.
The events are organized as a two-level hierarchy as follows:
An event log is made of a list of cases
(also known as traces in technical
term), and each case is made of a list of events. Cases and
events are both extensible constructs to allow for maximum flexibility in
event modeling.

The event logs provide a detailed record of the travel reimbursement process,
including important attributes such as identification, resource,
timestamp, permits, payment and budgeting information, and event labels.
Students or staff would submit their reimbursement application,
provide suppport documentation (if traveling internationally), wait for
approval and payment, and resubmit a corrected request if rejected.

As the logs were imported into Python via \textsc{pm4py},
they are represented as
dataframes, a key construct of
\textsc{pandas}, the popular Python Data Analysis Library. However, as
dataframes are not hierarchical, \textsc{pm4py} flattene the log so that 
the enclosing case and the case attributes
become attributes of an event. If
there were $A_c$ attributes associated with a
case, and $A_e$ event attributes in the
\textsc{xes} specification, after \textsc{pm4py} import, we would obtain a
\textsc{pandas} dataframe 
with $N$ events, each with a total of $A_c + A_e$ attributes.

In our project,
the domestic process log is organized differently than
the international process log.  For domestic trips, employees would first
complete the trip and then ask for reimbursement. On the other hand,
the policy for international trips mandates that
employees first obtain travel permits from the supervisor prior to
commencing the trips.

As a result, we are dealing with a more complex process in the case of
international travel, and the formal specification that describes
international cases also contain more fields.
Specifically, domestic cases have 5 attributes
each, and international cases have additional 13 attributes, together
with the 5 attributes shared with domestic cases
(\texttt{id}, \texttt{concept:name}, \texttt{BudgetNumber},
\texttt{DeclarationNumber}, \texttt{Amount}), for a total of 18 attributes.
Regardless of the type of trip,
the event in the \textsc{xes} hierarchy has the same 5 attributes
(\texttt{id}, \texttt{org:resource}, \texttt{concept:name},
\texttt{time:timestamp}, and \texttt{org:role}).
After flattening, \textsc{pm4py} returned a
domestic dataframe with $56437$ events and $5+5=10$ attributes, and
an international dataframe with $72151$ events and $18+5=23$ attributes.
To avoid namespace collision, all case attributes were prefixed with
\texttt{case:} by \textsc{pm4py} (for example, the \texttt{id} field of a case
became \texttt{case:id}). Please refer to Table
\ref{table-event} for the complete list of attributes:

@@

While we did not find any missing value per se (as far
as what \textsc{pandas} is concerned), we did find columns with string values
\textsc{unknown} or \textsc{missing}. We
carefully evaluated what to do with these attributes on a case-by-case
basis. For example, we could drop the column (attribute), drop the row
(event), fill the missing values, or simply ignore the column.

@@

\begin{figure*}[htbp]
\centerline{\includegraphics[width=0.55\textwidth]{images/domestic.png}}
\caption{\textsc{dfg} Model for Domestic Reimbursement.}
\label{fig-domestic}
\end{figure*}

\begin{figure*}[htbp]
\centerline{\includegraphics[width=0.55\textwidth]{images/international.png}}
\caption{\textsc{dfg} Model for International Reimbursement.}
\label{fig-international}
\end{figure*}

>>>>

\begin{abstract}
This project is a process mining case study designed to
identify patterns and
delays that obstruct the efficient flow of process
activities. Specifically, our
objective is to analyze event data logs from
the travel reimbursement process at Eindhoven University of Technology
(\textsc{tu/e}), identify bottlenecks in the process, and
seek improvement to its efficiency.
We prepare and filter the event data, and then generate simplified and
representative process models to locate
the most time-consuming and impactful critical tasks.
Our analysis suggests that
improvement in the supervisor approval and payment handling activities 
could significantly
reduce the time processing the travel reimbursement.
\end{abstract}

@@

Nowadays, most organizations use information systems to support
the execution of their business processes, generating significant
amounts of data.
Mainstream information systems, such as Enterprise Resource
Planning (\textsc{erp}), Work on Management Systems (\textsc{wms}), and Customer
Relationship Management (\textsc{crm}), have become vital tools behind the
operation of almost 
any company process\cite{Tuto2022}. All of these platforms
offer rich event logging capabilities, in which
an event log is
basically a record of an activity performed with a timestamp embedded
\cite{Proc2022}.

There are plenty of oppportunities, through
proper supervision of business behavior, to identify and adjust
tasks that are generating problems for the business processes
\cite{LeAl1994}.
One important technique towards such goal is process mining, a
specialized data analysis technique that reveals core business behavior
indicated by activity logs.
Process mining would start with these event logs, extracting important
information including the chronology of the activities
 and their execution time registered in each log. Representative models
would be built, or discovered automatically,
and then be used to detect errors and deficiencies in the process,
and to provide insight and suggestions for process enhancement.
Process mining is most closely related to data mining 
amongst business intelligence disciplines. However,
while data mining focuses on finding relationships among data,
process mining also looks at data from a systematic
perspective.

@@

In our process mining case study, we investigated
a dataset with
real-life event log data from the travel reimbursement process at
Eindhoven University of Technology (\textsc{tu/e}). Through analyzing the event
logs and mining the business process underlying the events, we seek to
identify and suggest remedies to address deficiencies in the reimbursement
process.
The event logs provide a detailed record of the travel reimbursement process,
including important attributes such as identification, resources,
timestamp, permits, payment and budgeting information, and event labels.
Students or staff would submit their reimbursement applications,
provide suppport documentation (especially
for international travels), wait for
reimbursement decision and subsequent payment, and possibly deal with
rejection and resubmission.
Collectively, these information provided a substantial amount of detail
into the reimbursement workflow. 
The data was obtained from 
the annual \textsc{bpi} Challenge as part of
the 2020 International Conference on Process Mining \cite{BPI2020}.

Our approach to process mining is based on the work by
Wil van der Aalst in his seminal paper titled simply
``Process Mining'' \cite{van2012},
while offered detailed process mining guidelines.
A process mining implementation generally
covers three main issues: (1) discovery, (2) conformance,
and (3) enhancement.
Our project focuses on discovery since it is the area
that requires the most effort and it is the most deterministic for the
success of the analysis. We will also take a cursory
look at conformance as both an aide to discovery and a motivation for
future work.

@@

Publications on practical application of process mining are scarce,
and our research paper aims to add support to this area. Our goal is
to demonstrate the practicality of process mining and analysis
in practice, in application to the real-life travel reimbursement problem
as detailed earlier.

Furthermore, we are very interested in the evaluation and application
of an emerging process mining approach based on
\textsc{pm4py}, a Python process mining library
developed by Berti et al.~\cite{BeZe2019}.
According to the authors,
\textsc{pm4py} is a modern, open-source, and extensible platform developed
specifically for the data science community. Prior to the turn of the
millenium,
there were hardly any tooling for process mining, but
since then a slew of software tools, both open-source
(\textsc{prom} and \textsc{apromore}) and commercial
(\textsc{disco}, \textsc{celonis}, and \textsc{processgold}),
have been developed for process mining.
Yet, many were proprietary, hard to extend,
or hidden behind graphical user interfaces.
The conservative mindset of process mining
tooling contrasted greatly with the modern philosophy of data science
R\&D,
which put significant emphasis on openness and extensibility.
\textsc{pm4py} was developed to bridge the gap
and bring process mining to mainstream data science
capabilities such as
\textsc{pandas}, \textsc{numpy}, \textsc{scikit-learn}.
We whole-heartedly
agree with the assessment of the authors, and look forward to opportunities
that the \textsc{pm4py} approach brought forth, including
the adoption of modern and rapidly evolving machine learning
techniques to process mining.

@@

This document is structured as follows: Sections \ref{section-intro} and
\ref{section-survey}
offer an introduction to process mining and a literature survey.
Sections \ref{section-technical}, \ref{section-methodology} and
\ref{section-models} describe the technical details of our methodology, as
well as the results of our investigation. We will end the paper with
future work in Section \ref{section-future} and conclusion in
\ref{section-conclusion}.

@@

\begin{figure}[htbp]
\centerline{\includegraphics[width=0.45\textwidth]{images/image2a.png}}
\caption{General outline of process mining approach.}
\label{fig-outline}
\end{figure}

@@

Many distinct data analysis approaches have been developed to discover
processes from event logs. However, \textsc{petri net}s\footnote{\textsc{petri net}s are also
equivalent to \textsc{bpmn} models and process trees.} have become a
widely adopted technique over other methods \cite{Cook1998} due to their
intuitiveness, mathematical soundness, and ease of analysis.
Another approach utilizes Unified
Modeling Language (\textsc{uml}) diagrams \cite{Sait2019}.
In 2004, van der Wilst et al.~described the 
a break-through algorithm in process mining
called \textsc{alpha miner}\cite{van2004} which
is capable of distilling event data logs into
work-flow diagrams, which can then be translated to \textsc{petri net}s.
The success of \textsc{alpha miner} has inspired many follow-up
publications,
including a popular model
called \textsc{heuristics miner} \cite{Weij2006}.
One main characteristic of this
algorithm is its high level of configurability and flexibility.
In order to deal with the
massive number of events that actual information systems generate,
Burattin et al.~\cite{Bura2014} developed another extension to
\textsc{heuristics miner} based on
lossy counting, as well as a sliding window that effectively analyzes
streams of real datasets.

@@

Most process mining publications are dedicated mainly to the enhancement
of novel methodologies and algorithms with significant emphasis on the
control-flow discovery area \cite{van2004}. However, the public services area
has been fruitful ground
for process mining deployment research. For instance, a
genetic miner approach was tested by de Madeiros et al.~\cite{deMe2007}
in Dutch municipalities. Additionally, van del Aalst et al.~\cite{RoDJ2009}
applied an analysis of organizations in another municipality process.
Rozinat et al.~\cite{RoMa2009} evaluated process models in two different case
studies on the public sector.

@@

We shall now briefly describe our approach to finding
where the process bottlenecks are. We would start with a thorough analysis and
cleaning of the event logs, and then
build models that best capture the behavior of the
process, making the proper tradeoff
between robustness (whether the model captures the intent of the process),
and coverage (what fraction of the traces are modeled).
 Then, we will use the
chosen models to answer the questions we hade in mind.

We shall note that process mining is a relatively new and unfamiliar
technique, and
as such we spent a disproportional amount of time ramping up the project.
While we completely believe in the full potential of \textsc{pm4py} due to
its focus on algorithmic discovery and availability
on Python, it is
essentially still an academic software in its nascent stage, and 
hence suffers from many issues such as software instability,
limitation in visualization and presentation, scanty documentation,
and immaturity in API design.  There is still plenty of room for
improvement.

@@

After a thorough review of what our tasks entail, we decided to focus
exclusively on the domestic and international travel event logs in order
to identify the process bottlenecks.  Both event logs
were provided to us in \textsc{xes} format,
an \textsc{xml}-based standard for event logs \cite{XES2021}, which can
be easily parsed using \textsc{pm4py}.
The events are organized as a two-level hierarchy as follows:
An event log is made of a list of cases
(also known as traces in research
terminology), and each case is made of a list of timestamped events. Cases and
events are both extensible constructs to allow for maximum flexibility in
event modeling.

The event logs provide a detailed record of the travel reimbursement process,
including important attributes such as identification, resources,
timestamp, permits, payment and budgeting information, and event labels.
Students or staff would submit their reimbursement application,
provide suppport documentation (if traveling internationally), wait for
approval and payment, and resubmit a corrected request if rejected.

The logs that were imported into Python via 
\textsc{pm4py} were converted into
dataframes, a key construct of
\textsc{pandas}, the popular Python Data Analysis Library. However, as
dataframes are not hierarchical, \textsc{pm4py} flattened the logs so that 
the enclosing case and the case attributes
became attributes of the event. If
there were $A_C$ attributes associated with a
case, and $A_E$ event attributes in the
\textsc{xes} specification, after \textsc{pm4py} import, we would obtain a
\textsc{pandas} dataframe 
with $N$ events, each with a total of $A_C + A_E$ attributes.

In our project,
the domestic process log is organized differently than
the international process log.  For domestic trips, employees would first
complete the trip and then ask for reimbursement. On the other hand,
the policy for international trips mandates that
employees first obtain travel permits from the supervisors prior to
commencing their trips.

As a result, the reimbursement process for international travel is
far more complex, and the formal specification that describes
international cases also contain more fields.
Specifically, while domestic cases only have 5 attributes
each, international cases have 13 additional attributes, on top of
the 5 attributes shared with domestic cases
(\texttt{id}, \texttt{concept:name}, \texttt{BudgetNumber},
\texttt{DeclarationNumber}, \texttt{Amount}), for a total of 18 attributes.
Regardless of the type of trips, however,
the event in the \textsc{xes} hierarchy has the same 5 attributes
(\texttt{id}, \texttt{org:resource}, \texttt{concept:name},
\texttt{time:timestamp}, and \texttt{org:role}).
After flattening, \textsc{pm4py} returned a
domestic dataframe with $56437$ events and $10~(=5+5$) attributes, and
an international dataframe with $72151$ events and $23~(=18+5$) attributes.
To avoid namespace collision, all case attributes were prefixed with
\texttt{case:} by \textsc{pm4py} (for example, the \texttt{id} field of a case
became \texttt{case:id}). Please refer to Table
\ref{table-event} for a complete list of all attributes:

@@

While we did not find any missing value per se (as far
as \textsc{pandas} is concerned), we did find columns with string values
\textsc{unknown} or \textsc{missing}. We
carefully evaluated what to do with these attributes on a case-by-case
basis. For example, we could drop the column (attribute), drop the row
(event), fill the missing values, or simply ignore the column.

@@

\begin{figure*}[htbp]
\centerline{\includegraphics[width=0.60\textwidth]{images/domestic.png}}
\caption{\textsc{dfg} Model for Domestic Reimbursement.}
\label{fig-domestic}
\end{figure*}
\vspace{-1em}

\begin{figure*}[htbp]
\centerline{\includegraphics[width=0.60\textwidth]{images/international.png}}
\caption{\textsc{dfg} Model for International Reimbursement.}
\label{fig-international}
\end{figure*}
\vspace{-1em}
